---
title: "EPID674 Epidemiologic Data Analysis using R"
subtitle: "Regression in R"
author: "Kelly Bakulski, Lauren Middleton"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

# Set up options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install new packages
```{r install_packages, include=FALSE, eval=FALSE}

# Install packages. Do this only once.
options(repos="https://cran.rstudio.com")
install.packages("tidyverse")
install.packages("here")
install.packages("ggpubr")
install.packages("broom")
install.packages("stargazer")
install.packages("sjlabelled")
# To avoid installing every time: change set up in curly brackets to eval=FALSE
```


# Load packages
```{r load_packages, include=FALSE}
##### Load these packages for the current session
library(tidyverse)
library(here)
library(ggpubr)
library(broom)
library(stargazer)
library(sjlabelled)
```


# Load data, recheck variables
```{r load_data}
# Check the file path
here("nhanes_dataset.rda")
# Load the saved R data
load(here("nhanes_dataset.rda"))

# Check the variables
str(nhanes_dataset)
```


# 6.1. Linear Models: Association between white blood cells (LBXWBCSI) and blood cadmium (LBXBCD)
## Outline
* Examine distributions of variables
* Check correlations between variables
* Check regression diagnostics
* Run linear regression



### Check cadmium variable first
```{r check_variables}

# Compare the distributions
bcd <- ggplot(nhanes_dataset,
              aes(x = LBXBCD)) +
  geom_histogram(fill = "darkorchid",
                 bins = 25)
bcd_log <- ggplot(nhanes_dataset,
              aes(x = log(LBXBCD))) +
  geom_histogram(fill = "seagreen3",
                 bins = 25)

# Does log-transformed cadmium look closer to the normal distribution?
ggarrange(bcd, bcd_log)
```

### Check white blood cell count variable
```{r check_variables}
# Check white blood cells
ggplot(nhanes_dataset,
              aes(x = LBXWBCSI)) +
  geom_histogram(bins = 25)

max(nhanes_dataset$LBXWBCSI, na.rm = TRUE)
# High measurement at 400 - NHANES says this is above upper limit of detection so we should remove it

# Remove the 400
nhanes_dataset <- nhanes_dataset %>%
  filter(!LBXWBCSI >= 400)
# Check new maximum value
max(nhanes_dataset$LBXWBCSI, na.rm = TRUE)

# Recheck white blood cells
ggplot(nhanes_dataset,
              aes(x = LBXWBCSI)) +
  geom_histogram(bins = 25)
```


## Look at bivariate association between log cadmium and continuous covariates
```{r}

# Set up the plots
age <- ggplot(nhanes_dataset,
              aes(x = RIDAGEYR,
                  y = log(LBXBCD))) +
  geom_point() +
  geom_smooth(method='lm')

pir <- ggplot(nhanes_dataset,
              aes(x = INDFMPIR,
                  y = log(LBXBCD))) +
  geom_point() +
  geom_smooth(method='lm')

wbc <- ggplot(nhanes_dataset,
              aes(x = log(LBXBCD),
                  y = LBXWBCSI)) +
  geom_point() +
  geom_smooth(method='lm')

# Plot the distributions
ggarrange(age, pir, wbc)
```


# Regression diagnostics: LBXBCD vs log(LBXBCD)
```{r linear_regression_diagnostics}
### Regression Diagnostics

# Set up the models for comparison. First is not log transformed
model <- lm(LBXWBCSI ~ LBXBCD,
                na.action = na.omit,
                data = nhanes_dataset)
model_log <- lm(LBXWBCSI ~ log(LBXBCD),
                na.action = na.omit,
                data = nhanes_dataset)


## In the case of linear model, the plot of the model gives diagnostic plots
# Residuals vs Fitted: linearity
# Scale-location: homogeneity of variance
# QQ: normality
# Residuals vs Leverage: outliers and leverage points

# non log-transformed model
par(mfrow = c(2, 2))
plot(model)

# log-transformed model
par(mfrow = c(2, 2))
plot(model_log)
```


# Simple linear regression
```{r simple_linear_regression}

# Create a simple regression model for white blood cell count and blood cadmium (crude model)
model_crude <- lm(LBXWBCSI ~ log(LBXBCD),
                  na.action = na.omit,
                  data = nhanes_dataset)

# print regression results
summary(model_crude)
# Is blood cadmium associated with white blood cell count?

# More detailed information about regression
summary.aov(model_crude)
glance(model_crude)

## Add age into the model
model_age <- lm(LBXWBCSI ~ log(LBXBCD) +
                  RIDAGEYR,
                na.action = na.omit,
                data = nhanes_dataset)
summary(model_age)
# Does anything change with age in the model?
```


# Add other variables to the linear regression
```{r linear_regression}

# Check that variables are factors, check the reference level
str(nhanes_dataset$race_eth) #5 levels for race with Non-Hispanic White as the reference
str(nhanes_dataset$sex) #2 levels for sex with Male as the reference

# Construct adjusted model
model_adj <- lm(LBXWBCSI ~ log(LBXBCD) +
                  RIDAGEYR +
                  race_eth +
                  sex +
                  INDFMPIR,
                na.action = na.omit,
                data = nhanes_dataset)
summary(model_adj)

# Why are there only 4 levels for race, not 5?
```


# Assign regression results to a dataframe
```{r save results as object}

# Model summaries
glance_model_adj <- as.data.frame(glance(model_adj))

# Model statistics
model_stats <- as.data.frame(tidy(model_adj))
```


# Make a nice table of regression statistics
  # For more information:
    *https://www.danieldsjoberg.com/gtsummary/articles/tbl_regression.html*
```{r gtsummary table}

tbl_regression(model_adj,
               label = c(race_eth ~ "Race/Ethnicity",
                         sex ~ "Sex")) %>%
  bold_labels()
```


# Compare regression models
```{r linear_regression_compare_models}

View(model_stats)
# What covariates are associated with white blood cells? Do they make sense biologically?

# Add other covariates to the model: which variables are biologically important?
# Set up models that add variables sequentially
m0 <- as.formula("LBXWBCSI ~ log(LBXBCD)")
m1 <- as.formula("LBXWBCSI ~ log(LBXBCD) + RIDAGEYR")
m2 <- as.formula("LBXWBCSI ~ log(LBXBCD) + RIDAGEYR + race_eth")
m3 <- as.formula("LBXWBCSI ~ log(LBXBCD) + RIDAGEYR + race_eth + sex")
m4 <- as.formula("LBXWBCSI ~ log(LBXBCD) + RIDAGEYR + race_eth + sex + INDFMPIR")

# Run the models and save as objects
model_0 <- lm(m0, na.action = na.omit, data = nhanes_dataset)
model_1 <- lm(m1, na.action = na.omit, data = nhanes_dataset)
model_2 <- lm(m2, na.action = na.omit, data = nhanes_dataset)
model_3 <- lm(m3, na.action = na.omit, data = nhanes_dataset)
model_4 <- lm(m4, na.action = na.omit, data = nhanes_dataset)

# Compare outputs from models side by side, save as html for viewing
stargazer(model_0, model_1, model_2, model_3, model_4,
          type = "html",
          dep.var.labels = "White Blood Cells (1000 cells/uL)",
          title = "Regression Results",
          out = "test.htm")
```


# Check confounders - 10% Guideline
```{r}
# Check if poverty-income ratio is a confounder

# Make the models into dataframes
model_no_pir <- as.data.frame(tidy(model_3))
model_pir <- as.data.frame(tidy(model_4)) #(same as model_stats)

# Pull out coefficients for blood cadmium from model 3 (unadjusted for poverty-income) and model 4, and turn the coefficients into vectors
est_model_no_pir <- model_no_pir %>%
  filter(term == "log(LBXBCD)") %>%
  pull(estimate)
est_model_pir <-  model_pir %>%
  filter(term == "log(LBXBCD)") %>%
  pull(estimate)

# 10% guideline for confounders: Does the addition of the new variable change the beta coefficient of interest by >10%?
# Calculate the percent change in the blood Pb coefficient before and after alcohol in the model
((est_model_pir - est_model_no_pir) / est_model_no_pir) * 100

# Does poverty-income ratio meet the guideline for a confounder?
```

# Effect modification by sex
```{r effect_modification}
# Does the relationship between blood cadmium and white blood cells vary by sex?

model_int <- lm(LBXWBCSI ~ log(LBXBCD) +
                  RIDAGEYR +
                  race_eth +
                  sex +
                  log(LBXBCD) * sex +
                  INDFMPIR,
                na.action = na.omit,
                data = nhanes_dataset)
model_int <- as.data.frame(tidy(model_int))

View(model_int)
# Is the interaction term significant?




### Stratify dataset by sex

# Check distribution
table(nhanes_dataset$sex)

# Filter dataset to only include male, run linear regression, save results as dataframe
model_male <- nhanes_dataset %>%
  filter(sex == "Male") %>%
  lm(LBXWBCSI ~ log(LBXBCD) +
                  RIDAGEYR +
                  race_eth +
                  INDFMPIR,
                na.action = na.omit,
                data = .) %>%
  tidy() %>%
  as.data.frame()
# What is the beta coefficient for blood Pb in males?

# Filter dataset to only include female, run linear regression, save results as dataframe
model_female <- nhanes_dataset %>%
  filter(sex == "Female") %>%
  lm(LBXWBCSI ~ log(LBXBCD) +
                  RIDAGEYR +
                  race_eth +
                  INDFMPIR,
                na.action = na.omit,
                data = .) %>%
  tidy() %>%
  as.data.frame()
# What is the beta coefficient for blood Pb in females? Is it similar to that in males?
```

########################################
############# Exercise 6A ##############
########################################

# Automate linear regressions with multiple outcomes
```{r}

# Goal: run individual linear regressions for 5 immune measure outcomes and 5 chemical exposures

# First create a long dataset from nhanes_dataset
    # Remove labels so columns can be vertically combined
    # Keep only the columns we want
    # Log transform the chemical measures
    # Make both the chemicals and immune measures long
    # Remove any rows with missing values

# Remove all labels
nhanes_unlab <- nhanes_dataset %>%
  remove_all_labels()

# Create vectors of the variables to keep
demog <- c("SEQN",
           "sex",
           "RIDAGEYR",
           "race_eth",
           "INDFMPIR",
           "education")
cells <- c("LBXRBCSI",
           "LBXWBCSI",
           "LBDLYMNO",
           "LBDNENO",
           "nlr")
chems <- c("LBXIRN",
           "URXUAS",
           "LBXCOT",
           "LBXBCD",
           "LBXBPB")

# Create the long dataset
long_nhanes_dataset <- nhanes_unlab %>%
  select(all_of(demog),
         all_of(cells),
         all_of(chems)) %>%
  mutate(LBXIRN = log(LBXIRN),
         LBXCOT = log(LBXCOT),
         URXUAS = log(URXUAS),
         LBXBCD = log(LBXBCD),
         LBXBPB = log(LBXBPB)) %>%
  pivot_longer(cols = all_of(chems),
               names_to = "chemical_codename",
               values_to = "chem_measurement") %>%
  pivot_longer(cols = all_of(cells),
               names_to = "celltype_codename",
               values_to = "cell_measurement") %>%
  na.omit()

# Run multiple linear regressions
 df_regressions_i <- long_nhanes_dataset %>%
    group_by(celltype_codename,
             chemical_codename) %>%
    do(lm(cell_measurement ~ chem_measurement +
                  RIDAGEYR +
                  race_eth +
                  sex +
                  INDFMPIR,
                  na.action = na.omit,
                  data = .) %>%
       tidy(.)) %>%
  ungroup(.)
 # How many regressions did we just run?
 # What is the formula for the first regression?
 
 View(model_stats_multiple)
```