---
title: "EPID674 Epidemiologic Data Analysis using R"
subtitle: "Homework 6B"
author: "Insert Your Name Here"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
output:
  word_document: default
---


```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(foreign)
library(epiDisplay)
library(stats)
library(Hmisc)
library(gmodels)

directory <- "/cloud/project/"
```

# The association between BPA and type-2 diabetes in a human population was reported first by Dr. David Melzer's group (Lang et al., JAMA 2008). Using NHANES 2003-2004 data, they found that a 1-SD increase in BPA was associated with a 39% increased odds of type-2 diabetes (OR=1.39, 95% CI, 1.21 to 1.60, p<0.001) after controlling for age, gender, race/ethnicity, education, income, smoking status, BMI, waist circumference, and urinary creatinine. Let’s try to examine the same research question using the expanded dataset that we have (NHANES 2003-2008, bpa.sas7bdat).


# Problem 1
## Construct your models as done by Lang: Model 1 adjusted for age, gender, and urinary creatinine; and Model 2 additionally adjusted for race/ethnicity, education, income, smoking, BMI, waist circumference. Compute odds ratios and 95% confidence intervals for a one standard deviation unit (1-SD) increase in BPA in each model. Interpret your results.
```{r problem 1}
#Insert code for problem 1 here
```
# Interpret results for problem 1 here
