---
title: "EPID674 Epidemiologic Data Analysis using R"
subtitle: "Homework 6A"
author: "Insert Your Name Here"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
output:
  word_document: default
---


```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(foreign)
library(epiDisplay)
library(stats)
library(Hmisc)
library(gmodels)
library(here)
```

# Using the dataset “bpa.rda”, answer the following. 


# Problem 1
## Build a regression model for the association between BMI and urinary BPA. Include urinary creatinine (ucr), age, gender, race-ethnicity, education and household income as important confounders. Check if smoking status is a confounder using a 10% rule. Based on your final model, what is the association between BMI and urinary BPA? Compute an effect estimate (difference in BMI) and 95% confidence intervals per one unit and IQR increase in urinary BPA. 

```{r problem 1}
#Insert code for problem 1 here
```
# Interpret results for problem 1 here

# Problem 2
## Verify the assumptions of a linear model for the final model. Create a histogram of the residuals and check if the distribution of residuals is skewed. Do you think log-transformation of BMI would improve the model fit?

```{r problem 2}
#Insert code for problem 2 here
```
# Include residuals histogram here
# Interpret results for problem 2 here

# Problem 3
# Check if gender modifies the association between urinary BPA and BMI. Which group is more susceptible?
```{r problem 3}
#Insert code for problem 3 here
```
# Interpret results for problem 3 here


