---
title: "EPID 674: Linear Regression in R"
subtitle: "Homework 6 Template"
author: "Insert your name here"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
output: 
  word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r load_packages, echo=T, results='hide', message=F, warning=F}

# Load packages for this assignment, do this every time
library(here)
library(ggcorrplot)
library(gtsummary)
library(epiDisplay)
library(tidyverse)
library(ggpubr)
library(broom)

```

# Load data
```{r read_data}

# Load saved NHANES homework dataset from the project files. This is the dataset we built in homework 2
load(here(("nhanes_homework_dataset.rda")), verbose = TRUE)
```

# 1. 1.	Prior to using linear regression, develop an understanding of the bivariate relationship between your exposure and outcome.



```{r bivariate_assessment}

# Insert code for problem 1 here

```
Interpret the findings: 


# 2. Use simple linear regression to test the unadjusted association between blood cadmium levels (variable: LBXBCD) and cognition score (CFDDS)

```{r cadmium_cognition}

# Insert code for problem 2 here

```

Interpret the findings: 

# 3. Use simple linear regression to test the unadjusted association between educational attainment (variable: education) and cognition score (variable: CFDDS).


```{r education_cognition}

# Insert code for problem 3 here

```

Interpret the findings: 




# 4. Use multivariable linear regression to test the adjusted association between blood cadmium levels (variable: LBXBCD) and cognition score (variable: CFDDS), after adjusting for educational attainment, sex, race/ethnicity, and age.

```{r multivariable}

# Insert code for problem 4 here

```

Interpret the findings: 


