---
title: "EPID674 Epidemiologic Data Analysis using R"
subtitle: "Regression in R"
author: "Kelly Bakulski"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
output:
   word_document: default
   html_document: 
    toc: true
    toc_float: true
---

# Set up options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install new packages
```{r install packages, eval=FALSE}

# Install packages. Do this only once.
options(repos="https://cran.rstudio.com" )
install.packages("epiDisplay")
install.packages("Hmisc")
install.packages("foreign")
install.packages("stargazer")
# To avoid installing every time: change set up in curly brackets to eval=FALSE
```


# Load packages
```{r load packages, include=FALSE}
##### Load these packages for the current session
library(foreign)
library(Hmisc)
library(epiDisplay)
library(stargazer)
```


# Specify file directories
```{r directories}
#directory <- "/cloud/project/" #Class option when coding on RStudio Cloud

# Three alternate options to customize and specify source data directory if programming on a personal computer (only need to do the one that works for you)
# directory<-"M:/EPID674/Data/" # Alternate option 1: location of the parent dataset in afs space of school computer
 directory <- "/Volumes/GoogleDrive/My Drive/Teaching/EPID674/2020_fall/EPID674_Week5_Class/" # Alternate option 2: location of data on my Macbook
# directory<-"C:/Users/bakulski/Documents/EPID674/Data/" #Alternate option 3: location of data on a personal PC

date <- format(Sys.Date(), "%Y%m%d")
```


# Load data, remake useful variables
```{r read data}

# Check the file path
paste0(directory, "nhanes3.rda")
# Load the saved R data
load(paste0(directory, "nhanes3.rda"))
### Or load RData if you saved the workspace image
# load(".RData")


# Remake a few variables from last class if they are no longer in your environment
sex1 <- factor(nhanes$sex, levels = c(1, 2), labels = c("male", "female"))
AGE5b <- cut(nhanes$age, quantile(nhanes$age, c(0, .2, .4, .6, .8, 1)), include.lowest = T) # quintiles
AGE5c <- cut(nhanes$age, breaks = c(19, 40, 50, 60, 70, 90))
age5c <- unclass(AGE5c)
nhanes <- cbind(nhanes, sex1, AGE5b, AGE5c, age5c)
```


### 6.1. Linear Models: Association between systolic blood pressure (SBP) and blood lead (bpb)
```{r check variables}
## Does the distribution of log(sbp) look closer to the normal distribution?
par(mfrow = c(2, 1))
hist(nhanes$sbp, nclass = 20, col = "darkorchid")
hist(log(nhanes$sbp), nclass = 20, col = "seagreen3")

## Let's start with non-log transformed SBP first. Look at bivariate association between sbp and continuous covariates
par(mfrow = c(1, 1))
plot(nhanes$age, nhanes$sbp, pch = 20, cex = 0.7, col = "dodgerblue", cex.lab = 1.2, las = 1, ylab = "SBP", xlab = "Age (years)")
lines(smooth.spline(nhanes$age, nhanes$sbp, df = 10), col = "dodgerblue", lwd = 3)
plot(nhanes$bmi, nhanes$sbp, pch = 20, cex = 0.7, col = "dodgerblue", cex.lab = 1.2, las = 1, ylab = "SBP", xlab = "BMI (kg/m2)")
lines(smooth.spline(na.omit(nhanes$bmi), nhanes$sbp[na.omit(nhanes$bmi)], df = 3, ), col = "grey60", lwd = 3)
plot(nhanes$bpb, nhanes$sbp, pch = 20, cex = 0.7, col = "dodgerblue", cex.lab = 1.2, las = 1, ylab = "SBP", xlab = "Blood lead level (ug/dL)")
lines(smooth.spline(nhanes$bpb, nhanes$sbp, df = 10), col = "grey60", lwd = 3)

# Clean up dataset. Make sure NaN are NA for key covariates
table(nhanes$educ, useNA = "always")
nhanes$educ[is.nan(nhanes$educ)] <- NA
table(nhanes$educ, useNA = "always")
table(nhanes$alc, useNA = "always")
nhanes$alc[is.nan(nhanes$alc)] <- NA
table(nhanes$alc, useNA = "always")
```


# Simple linear regression
```{r simple linear regression}
## Start creating a simple regression model for systolic blood pressure,
## including only blood lead (bpb) (crude model).
sbp.model <- lm(sbp ~ bpb, na.action = na.omit, data = nhanes)
sbp.model
summary(sbp.model)
summary.aov(sbp.model)
anova(sbp.model)

## Add age in the model
sbp.model1 <- lm(sbp ~ bpb + age, na.action = na.omit, data = nhanes)
summary(sbp.model1)
```

# Add categorical variables linear regression
```{r linear regression categorical variables}
## Add race, which is a categorical variable, therefore we should use factor()
## or as.factor() which creates indicator variables for each category of race
table(nhanes$race)
sbp.model2 <- lm(sbp ~ bpb + age + factor(race), na.action = na.omit, data = nhanes)
summary(sbp.model2)

#### Change the reference level for a factor variable
# Method 1: Make a new (or alter the old) variable
table(nhanes$race)
race2 <- relevel(factor(nhanes$race), ref = 2)
table(race2)
sbp.model2 <- lm(sbp ~ bpb + age + factor(race2), na.action = na.omit, data = nhanes)
summary(sbp.model2)
# Method 2: Change the contrasts in the lm() statement
sbp.model2 <- lm(sbp ~ bpb + age + C(factor(race), contr.treatment(2, base = 2)), na.action = na.omit, data = nhanes)
summary(sbp.model2)

# R provides Type I sequential SS, not the default Type III marginal SS reported by SAS and SPSS.
# We will need to use the drop1() function to produce the familiar Type III results.
# It will compare each term with the full model.
anova(sbp.model2)
drop1(sbp.model2, test = "F")

## alternative
library(car)
Anova(sbp.model2, type = "III")

## We can also use the update function to add variables to a pre-existing model
sbp.model2 <- update(sbp.model1, . ~ . + factor(race))
summary(sbp.model2)
```

# Compare regression models
```{r linear regression compare models}
## Add other covariates to the model: which variables are biologically important?
## Add sex, BMI, educ and smk.
sbp.model3 <- lm(sbp ~ bpb + age + factor(race) + factor(sex) + bmi + factor(educ) + factor(smk), na.action = na.omit, data = nhanes)
summary(sbp.model3) # What covariates are associated with SBP? Do they make sense biologically?

# See output from models 1,2,and 3 side by side
stargazer(sbp.model1, sbp.model2, sbp.model3, type = "text", dep.var.labels = "Systolic Blood Pressure (mmHg)")

### Check if alcohol consumption is a confounder
sbp.model4 <- update(sbp.model3, . ~ . + factor(alc))
summary(sbp.model4) # Is alcohol associated with sbp?

summary(sbp.model4)$coef #  Pull out relevant information from the output: Coefficient table
summary(sbp.model4)$coef[2, 1] # Pull out just the beta coefficient blood Pb
# 10% guideline for confounders: Does the addition of the new variable change the beta coefficient of interest by >10%?
(summary(sbp.model4)$coef[2, 1] - summary(sbp.model3)$coef[2, 1]) / summary(sbp.model3)$coef[2, 1] # Calculate the percent change in the blood Pb coefficient before and after alcohol in the model
# Does alcohol meet the guideline for a confounder?
```

# Reporting regression results
```{r linear regression reporting results}
## Compute percent increase in SBP and 95% CI per one unit increase and IQR increase in bpb
# for one unit increase in exposure
summary(sbp.model4)$coef[2, 1] # Beta coefficent
summary(sbp.model4)$coef[2, 1] - 1.96 * summary(sbp.model4)$coef[2, 2] # Lower confidence interval
summary(sbp.model4)$coef[2, 1] + 1.96 * summary(sbp.model4)$coef[2, 2] # Upper confidence interval
regress.display(sbp.model4) # Alternative: Convenience function to show all of the CI with the beta coefficients

# for an IQR increase in exposure, to improve communication of findings
IQR(nhanes$bpb)
IQR(nhanes$bpb) * summary(sbp.model4)$coef[2, 1] # Per IQR increase in exposure, beta coefficent 
l95ci.iqr <- IQR(nhanes$bpb) * (summary(sbp.model4)$coef[2, 1] - 1.96 * summary(sbp.model4)$coef[2, 2]) # Per IQR increase in exposure, lower confidence interval
u95ci.iqr <- IQR(nhanes$bpb) * (summary(sbp.model4)$coef[2, 1] + 1.96 * summary(sbp.model4)$coef[2, 2]) # Per IQR increase in exposure, upper confidence interval
regress.display(sbp.model4)$table[2, ] * IQR(nhanes$bpb) # Multiply the convenience table by the IQR
```

# Compare regression results
```{r linear regression compare results}
## Does the variable packyrs give a better estimate than smk?
sbp.model5 <- update(sbp.model4, . ~ . - factor(smk) + packyrs)
summary(sbp.model5)

AIC(sbp.model4, sbp.model5) # Which model has the lower AIC? The one with smk (model 4) or the one with packyrs (model 5)?

## Run two models one with age and one adding age as a quadratic function.
## Does the quadratic term improve the fit of the model?
sbp.model6 <- update(sbp.model4, . ~ . + I(age^2))
summary(sbp.model6) # Is the age squared term significant?

## use the anova function to compare the 2 models and
## see if the quadratic term improves the model.
anova(sbp.model4, sbp.model6, test = "F")

## so sbp.model6 is our final model
## What is the relationship between bpb and sbp?
```

# Regression diagnostics
```{r linear regression diagnostics}
### Regression Diagnostics
## In the case of linear model, the plot of the model gives diagnostic plots
par(mfrow = c(1, 1))
plot(sbp.model6)

par(mfrow = c(2, 2))
plot(sbp.model6, id.n = 5, cex = 0.1)
# plot.lm(sbp.model6)

par(mfrow = c(1, 1))
plot(sbp.model6, which = 1) # Can ask for the four plots one at a time
plot(sbp.model6, which = 2)
plot(sbp.model6, which = 3)
plot(sbp.model6, which = 5)
plot(sbp.model6, which = 4) # There is an extra, hidden plot you can call out individually

## How about log(sbp)?
## Check how diagnostic plots using log-transformed sbp look like

sbp.model6.log <- update(sbp.model6, . - sbp + log(sbp) ~ .)
summary(sbp.model6.log)
plot(sbp.model6.log)
hist(residuals(sbp.model6.log), nclass = 20)

par(mfrow = c(2, 2))
plot(sbp.model6, which = 1)
plot(sbp.model6.log, which = 1)
hist(residuals(sbp.model6), main = "Histogram of res(SBP)")
hist(residuals(sbp.model6.log), main = "Histogram of res(log(SBP))")
par(mfrow = c(1, 1))

### Partial Residual Plots
## Plot sbp vs. bpb given that other variables are in the model (adjusted)

## This can be done by 'termplot'
termplot(sbp.model6, partial.resid = TRUE, col.res = "gray30", cex = 0.3, smooth = panel.smooth)
```

## Suppose you decide to go with log transformed SBP.
## Compute percent increase in SBP and 95% CI per one unit increase and IQR increase in bpb
## what to report if y is log-transformed?
## percent increase (difference) vs. absolute increase (difference)?
## Compute percent increase in SBP and 95% CI per one unit increase and IQR increase in bpb
```{r log transformed outcome reporting results}
summary(sbp.model6.log)$coef
summary(sbp.model6.log)$coef[2, 1]
summary(sbp.model6.log)$coef[2, 2]
exp(summary(sbp.model6.log)$coef[2, 1])
# exp(sbp.model6.log$coef[2])
100 * (exp(summary(sbp.model6.log)$coef[2, 1]) - 1) # Per one unit increase in blood Pb, percent change in systolic blood pressure
100 * (exp(summary(sbp.model6.log)$coef[2, 1] - 1.96 * summary(sbp.model6.log)$coef[2, 2]) - 1) # Lower confidence interval
100 * (exp(summary(sbp.model6.log)$coef[2, 1] + 1.96 * summary(sbp.model6.log)$coef[2, 2]) - 1) # Upper confidence interval


IQR(nhanes$bpb)
100 * (exp(IQR(nhanes$bpb) * summary(sbp.model6.log)$coef[2, 1]) - 1) # Per IQR unit increase in blood Pb, percent change in systolic blood pressure
100 * (exp(IQR(nhanes$bpb) * (summary(sbp.model6.log)$coef[2, 1] - 1.96 * summary(sbp.model6.log)$coef[2, 2])) - 1) # Lower confidence interval
100 * (exp(IQR(nhanes$bpb) * (summary(sbp.model6.log)$coef[2, 1] + 1.96 * summary(sbp.model6.log)$coef[2, 2])) - 1) # Upper confidence interval
```
## linearity assumption

# Non-linear associations
```{r add a smoothing term}
# What if the relationship between blood lead and sbp is non-linear? 
library(mgcv)

sbp.model6.gam <- gam(sbp ~ s(bpb) + age + factor(race) + factor(sex) + bmi + factor(educ) + factor(smk) + factor(alc) + I(age^2), na.action = na.omit, data = nhanes)
summary(sbp.model6.gam)
plot(sbp.model6.gam)
plot(sbp.model6.gam, xlab = "Blood lead (ug/dL)", ylab = "Change in log(SBP)") # Does this relationship look linear?

# Option to log transform the exposure
sbp.model7 <- lm(sbp ~ log(bpb) + age + factor(race) + factor(sex) + bmi + factor(educ) + factor(smk) + factor(alc) + I(age^2), na.action = na.omit, data = nhanes)
summary(sbp.model7)
summary(sbp.model6)

## compare model6 and model7
AIC(sbp.model6, sbp.model7) # Which model has the lower AIC? The spline term or the log term for blood Pb?
```

# Effect modification by sex
```{r effect modification}
# What if the relationship between blood Pb and sbp varies by sex?
table(nhanes$sex)

### Stratified by sex
## Male (sex==1)
sbp.model6.male <- lm(sbp ~ bpb + age + factor(race) + bmi + factor(educ)
  + factor(smk) + factor(alc) + I(age^2), data = nhanes, subset = (sex == 1))
summary(sbp.model6.male) # What is the beta coefficent for blood Pb in males?

## Female (sex==2)
sbp.model6.female <- lm(sbp ~ bpb + age + factor(race) + bmi + factor(educ)
  + factor(smk) + factor(alc) + I(age^2), data = nhanes, subset = (sex == 2))
summary(sbp.model6.female) # What is the beta coefficent for blood Pb in females? Is it similar to that in males?

### Use interaction model
sbp.model6.int <- lm(sbp ~ bpb + factor(sex) + bpb * factor(sex) + age + factor(race) + bmi + factor(educ)
  + factor(smk) + factor(alc) + I(age^2), data = nhanes)
# below is the same
sbp.model6.int <- lm(sbp ~ bpb * factor(sex) + age + factor(race) + bmi + factor(educ)
  + factor(smk) + factor(alc) + I(age^2), data = nhanes)
summary(sbp.model6.int) # Is the interaction term significant?
```
########################################
############ Exercise 6A ##############
########################################


### 6.3. Generalized Linear Models: Association between hypertension (htn) and blood lead (bpb)
```{r logistic regression}
## Logistic regression for hypertension
## Look at hypertension status (htn)
tab1(nhanes$htn, graph = F)

htn.model <- glm(htn ~ bpb + age + factor(sex) + factor(race) + bmi + factor(educ) + factor(smk) + factor(alc),
  family = binomial,
  na.action = na.omit, data = nhanes
)
summary(htn.model)

# Compute ORs
logistic.display(htn.model)

# Regression diagnostics
par(mfrow = c(2, 2))
plot(htn.model)

par(mfrow = c(1, 1))
termplot(htn.model, se = T, partial.resid = T)
```

# Stepwise variable selection
```{r logistic regression stepwise variable selection, include = F}
## you can construct your model using variable selection methods (forward, backward, stepwise)
# first, define a matrix of predictors
X <- nhanes[, c(3:8, 13:25, 29, 30)]
Y <- nhanes$htn
dat <- data.frame(cbind(Y, X))
dat <- na.omit(dat) # note: dataset should be complete
summ(dat)

fit.start <- glm(Y ~ 1, data = dat, family = binomial) # Starting model has no predictors, just outcome.
summary(fit.start)
fit.full <- glm(Y ~ ., data = dat, family = binomial) # Full model would have all possible covariates
summary(fit.full)

# Step forward from an empty model to select the most informative predictors
full <- formula(glm(Y ~ ., data = dat, family = binomial))
full
fit.forward <- step(fit.start, direction = "forward", scope = full)
summary(fit.forward) # Is blood Pb in the final model?

# Step backward from a full model to select the informative predictors (always keep bpb in the model)
fit.backward <- step(fit.full, direction = "backward", scope = list(lower = ~bpb))
summary(fit.backward) # Is blood Pb significant?

# Step forward and backward (always keep bpb in the model)
fit.step <- step(fit.full, direction = "both", scope = list(lower = ~bpb))
summary(fit.step) # This is exactly the same as backward
```

# Poisson Regression for count outcomes. 
# Code is available, but we will skip going through this during class
```{r poisson regression}
# ## Poisson regression for respiratory death: Montana dataset from epiDisplay
# data(Montana)
# summ(Montana)
# head(Montana, 10)
# hist(Montana$respdeath)
# 
# par(mfrow = c(2, 2))
# tab1(Montana$agegr)
# tab1(Montana$period)
# tab1(Montana$start)
# tab1(Montana$arsenic)
# 
# ## label categorical variables
# Montana$agegr <- factor(Montana$agegr, labels = c("40-49", "50-59", "60-69", "70-79"))
# Montana$period <- factor(Montana$period, labels = c("1938-1949", "1950-1959", "1960-1969", "1970-1977"))
# Montana$start <- factor(Montana$start, labels = c("pre-1925", "1925 & after"))
# Montana$arsenic <- factor(Montana$arsenic, labels = c("<1 year", "1-4 years", "5-14 years", "15+ years"))
# 
# tab1(Montana$agegr, missing = F)
# tab1(Montana$period, missing = F)
# tab1(Montana$start, missing = F)
# tab1(Montana$arsenic, missing = F)
# par(mfrow = c(1, 1))
# 
# ## Compute incidence rate by age and period
# table.pyears <- tapply(Montana$personyrs, list(Montana$period, Montana$agegr), sum)
# table.deaths <- tapply(Montana$respdeath, list(Montana$period, Montana$agegr), sum)
# table.inc10000 <- table.deaths / table.pyears * 10000
# table.inc10000
# 
# ## create a time-series plot of the incidence
# plot.ts(table.inc10000, plot.type = "single", xlab = "", ylab = "#/10,000 person-years", xaxt = "n", col = c("black", "blue", "red", "green"), lty = c(2, 1, 1, 2), las = 1)
# points(rep(1:4, 4), table.inc10000, pch = 22, cex = table.pyears / sum(table.pyears) * 20)
# title(main = "Incidence by age and period")
# axis(side = 1, at = 1:4, labels = levels(Montana$period))
# legend("topleft", legend = levels(Montana$agegr)[4:1], col = c("green", "red", "blue", "black"), bg = "white", lty = c(2, 1, 1, 2))
# 
# ## check arsenic
# tab1(Montana$arsenic)
# tapply(Montana$respdeath, Montana$arsenic, mean)
# tapply(Montana$personyrs, Montana$arsenic, mean)
# 
# ## Poisson model
# resp.mode11 <- glm(respdeath ~ period, offset = log(personyrs), family = poisson, data = Montana)
# summary(resp.mode11)
# 
# resp.mode12 <- glm(respdeath ~ agegr, offset = log(personyrs), family = poisson, data = Montana)
# summary(resp.mode12)
# 
# resp.mode13 <- glm(respdeath ~ period + agegr, offset = log(personyrs), family = poisson, data = Montana)
# summary(resp.mode13)
# 
# AIC(resp.mode11, resp.mode12, resp.mode13)
# ## model2 is better
# 
# resp.mode14 <- glm(respdeath ~ agegr + arsenic, offset = log(personyrs), family = poisson, data = Montana)
# summary(resp.mode14)
# 
# ## is there a linear trend across arsenic exposure?
# resp.mode14.lin <- glm(respdeath ~ agegr + as.numeric(arsenic), offset = log(personyrs), family = poisson, data = Montana)
# summary(resp.mode14.lin)
# 
# ## compute IRR
# idr.display(resp.mode14)
```

########################################
############ Exercise 6B ##############
########################################

### 6.5. Matched Case-Control Study: VC1to1 from epiDisplay
```{r matched case-control}
# Matched case-control dataset available in the epiDisplay package
data(VC1to1)
summ(VC1to1) # What is the shape/structure of this dataset?
head(VC1to1) # 1 case matched to 1 control

# Reshape the data to facilitate data exploration
# function 'reshape' converts wide to long or vice versa
wide <- reshape(VC1to1, timevar = "case", v.names = c("smoking", "rubber", "alcohol"), idvar = "matset", direction = "wide")
head(wide, 3)
table(wide$smoking.1, wide$smoking.0, dnn = c("smoking in case", "smoking in control"))
# dnn: dimnames names

# matchTab() is for the conditional OR (McNemar's OR)
matchTab(VC1to1$case, VC1to1$smoking, strata = VC1to1$matset) # Is smoking exposure associated with cancer?
matchTab(VC1to1$case, VC1to1$rubber, strata = VC1to1$matset) # Is working in the rubber industry associated with cancer?
matchTab(VC1to1$case, VC1to1$alcohol, strata = VC1to1$matset) # Is alcohol consumption associated iwth cancer?

## look at the full dataset VC1to6
data(VC1to6) # 1 case matched to up to 6 controls
summ(VC1to6)
VC1to6[, ]

# what is the effect of smoking?
matchTab(VC1to6$case, VC1to6$smoking, strata = VC1to6$matset) # Is smoking exposure associated with cancer?
matchTab(VC1to6$case, VC1to6$alcohol, strata = VC1to6$matset)
```

# Conditional logistic regression
```{r conditional logistic regression}
# conditional logistic reg using clogit from survival package
### load the survival package
library(survival)
clogit1 <- clogit(case ~ smoking + alcohol + strata(matset), data = VC1to1) # 1 to 1 match dataset
summary(clogit1) # What covariates are associated?

clogit2 <- clogit(case ~ smoking + alcohol + strata(matset), data = VC1to6) # 1 to 6 match dataset
summary(clogit2) # What covariates are associated?

# compute ORs
clogistic.display(clogit1)
clogistic.display(clogit2)
```

### 6.6. Survival Analysis: Association between total mortality (d_total) and blood lead (bpb)
```{r survival analysis}
tab1(nhanes$d_total) # Variable for death due to any cause in NHANES
summ(nhanes$pmon_mec) # Number of months of follow up 

### Define Surv() object to use in later functions
surv.total <- Surv(nhanes$pmon_mec, nhanes$d_total)
surv.total[1:10] # Includes information on time of follow up and whether the person died, denoted with a "+"

### K-M Life table and curve
fit.total <- survfit(Surv(nhanes$pmon_mec, nhanes$d_total) ~ 1) # Model with no predictors, just outcome
#summary(fit.total)
summary(nhanes$pmon_mec/12) # How many years of follow up do you have?
fit.total # How many people died over this time period?
plot(fit.total)
## suppress 95% CI lines and the time marks for censored subjects.
plot(fit.total, ylim = c(0.7, 1.0), conf.int = F, mark.time = F, ylab="Probability of survival", xlab="Months")

### Survival by different levels of covariates
fit.total.sex <- survfit(Surv(nhanes$pmon_mec, nhanes$d_total) ~ nhanes$sex)
fit.total.sex # How many people died in each sex group?
#summary(fit.total.sex)[1:10]

plot(fit.total.sex, ylim = c(0.6, 1.0), col = c("blue", "red"), lty = c(1, 2), mark.time = F, main = "Kaplan-Meier curve", xlab = "Time (months)", ylab = "Survival probability")
legend("bottomleft", legend = c("Men", "Women"), col = c("blue", "red"),  lty = c(1, 2))

### Test for differences in survival curves
survdiff(Surv(nhanes$pmon_mec, nhanes$d_total) ~ nhanes$sex) # Is there a difference in survival among males and females?
```

### Cox regression
```{r cox regression}
cox.bpb <- coxph(Surv(nhanes$pmon_mec, nhanes$d_total) ~ nhanes$bpb)
summary(cox.bpb) # Is blood Pb associated with survival? Easier to visualize in categories of exposure

bpb3 <- cut2(nhanes$bpb, g = 3)
tab1(bpb3) # Tertiles of blood Pb

# K-M Life table and curve
fit.total.bpb3 <- survfit(Surv(nhanes$pmon_mec, nhanes$d_total) ~ bpb3)
fit.total.bpb3
plot(fit.total.bpb3, col = c(1:3), lty = c(1:3), ylim = c(0.6, 1.0), main = "Survival in relation to blood lead levels", xlab = "Time (months)", ylab = "Survival probability") 
legend(30, 0.7, legend = c("Q1", "Q2", "Q3"), lty = c(1:3), col = c(1:3))

# crude
cox.bpb3 <- coxph(Surv(nhanes$pmon_mec, nhanes$d_total) ~ bpb3)
summary(cox.bpb3)

# adjusted
cox.bpb3.adj <- coxph(Surv(nhanes$pmon_mec, nhanes$d_total) ~ bpb3 + nhanes$age + factor(nhanes$sex) + factor(nhanes$race) + factor(nhanes$educ) + factor(nhanes$smk) + factor(nhanes$alc))
summary(cox.bpb3.adj)

### Test for the proportional hazards assumption
test.prop <- cox.zph(cox.bpb3.adj) # Do any of the variables violate the proportional hazards assumption? May consider stratifying by sex.
test.prop 
## Display a graph of the scaled Schoenfeld residuals, along with a smooth curve
plot(test.prop) # for all variables
plot(test.prop, var = 4) + # Can call up variables indivdually, here for sex
abline(h = 0, lty = 3, col = 2)

# Stratify by sex
cox.bpb3.adj1 <- coxph(Surv(nhanes$pmon_mec, nhanes$d_total) ~ bpb3 + nhanes$age + strata(nhanes$sex) + factor(nhanes$race) + factor(nhanes$educ) + factor(nhanes$smk) + factor(nhanes$alc))
summary(cox.bpb3.adj1)
cox.zph(cox.bpb3.adj1)
```

### 6.7. Write your own functions
```{r write your own simple function}
## Let's make a simple macro that calculates the mean and standard deviation at the same time.

# get the results in vector form
mystats <- function(x) {
  mymean <- mean(x, na.rm = T)
  mysd <- sd(x, na.rm = T)
  c(mean = mymean, sd = mysd)
}

mystats(nhanes$age)
mystats(nhanes$bpb)
```

# Regression functions 
```{r write your own regression function}
# Assume that you are examining age-adjusted associations of SBP with from the 13th variables (bmi) to 25th variables (packyrs) (n=13)
# you want to run 13 linear regression models and save beta's and p-values

summ(nhanes)
test.var <- nhanes[, c(13:25)]
head(test.var)

# example
mod <- lm(sbp ~ bmi + age, data = nhanes, na.action = na.omit)
summary(mod)
summary(mod)$coef[2, 1] # how to extract beta for bmi
summary(mod)$coef[2, 4] # how to extract p-value for bmi

Test <- function(data, y, cov) {
  nvar <- ncol(data)
  newdata <- data.frame(cbind(data, y, cov))

  tmatrix <- data.frame(matrix(NA, 2, nvar)) # 2 rows
  colnames(tmatrix) <- colnames(data) # create a row for each var
  rownames(tmatrix) <- c("beta", "p")

  for (i in 1:nvar) {
    ind <- data[, i]
    model <- lm(y ~ ind + cov, data = newdata, na.action = na.omit)
    tmatrix[1, i] <- summary(model)$coef[2, 1]
    tmatrix[2, i] <- summary(model)$coef[2, 4]
  }
  return(tmatrix)
  write.csv(tmatrix, file = paste0(output_dir, "sbp.results.csv"))
}

Test(test.var, nhanes$sbp, nhanes$age) # Look in your output directory for the beta coeficients and p-values for all covariates! 
```

########################################
############ Optional: Exercise 6C ##############
########################################
